---
output:
  pdf_document: default
  html_document: default
---
# Rejection Sampling

## Overview
If $f(x)$ can be calculated, then we can use rejection sampling to obtain a random draw from exactly the target distribution. This strategy relies on sampling candidates from an easier distribution and then correcting the sampling probability through random rejection of some candidates.

Let $g$ denote another density from which we know how to sample and for which we can easily calculate $g(x)$. Let $e(\cdot)$ denote an envelope, having the property $e(x)=$ $g(x) / \alpha \geq f(x)$ for all $x$ for which $f(x)>0$ for a given constant $\alpha \leq 1$. Rejection sampling proceeds as follows:   
**1.** Sample $Y \sim g$.   
**2.** Sample $U \sim \operatorname{Unif}(0,1)$.   
**3.** Reject $Y$ if $U>f(Y) / e(Y)$. In this case, do not record the value of $Y$ as an element in the target random sample. Instead, return to step 1 .    
**4.** Otherwise, keep the value of $Y$. Set $X=Y$, and consider $X$ to be an element of the target random sample. Return to step 1 until you have accumulated a sample of the desired size.   

The draws kept using this algorithm constitute an i.i.d. sample from the target density $f$; there is no approximation involved. To see this, note that the probability that a kept draw falls at or below a value $y$ is
$$
\begin{aligned}
P[X \leq y] &=P\left[Y \leq y \mid U \leq \frac{f(Y)}{e(Y)}\right] \\
&=P\left[Y \leq y \text { and } U \leq \frac{f(Y)}{e(Y)}\right] / P\left[U \leq \frac{f(Y)}{e(Y)}\right] \\
&=\int_{-\infty}^y \int_0^{f(z) / e(z)} d u g(z) d z / \int_{-\infty}^{\infty} \int_0^{f(z) / e(z)} d u g(z) d z \\
&=\int_{-\infty}^y f(z) d z
\end{aligned}
$$
which is the desired probability.   

## Example
Suppose we want to estimate $S=E[X^2]$ when X has the density that is proportional to $q(x) = exp\{-|x|^3/3\}$. Use rejection sampling to estimate $E[X^2]$. 
Let g denote standard normal(Gaussian) distribution, which we know how to sample and $g(x)$ can be easily calculated. Let $e(\cdot)$ denote an envelope, having property that $e(x)=g(x)/\alpha\geq f(x)$.  
First, we want to find $\alpha$ and $e(\cdot)$. To find $\alpha$ such that $g(x)/\alpha\geq f(x)$ for all x.  
$$exp\{-x^2/2\}/\sqrt{2\pi}/\alpha\geq exp\{-|x|^3/3\}$$
$$1/\alpha\geq exp\{-|x|^3/3+x^2/2\}\sqrt{2\pi}$$
$$1/\alpha\geq exp\{(3x^2-2|x|^3)/6\}\sqrt{2\pi}$$
$$1/\alpha = 2.961$$
$$e(x)=g(x)*2.961$$
Code f(x) in R:   
```{r}
f_x <- function(x) {
  exp(-abs(x)^3/3)
}
```
Code e(x) in R:   
```{r}
e_x <- function(x) {
  2.961*exp(-x^2/2)/sqrt(2*pi)
}
```
Step 1: Sample $Y\sim g$   
```{r}
set.seed(1693)
N<- 10^6
Y<- rnorm(N)
```
Step 2: Sample $U\sim Unif(0,1)$   
```{r}
U<- runif(N)
```
Step 3 and 4: Reject Y if $U > f(Y)/e(Y)$, Record X=Y if $U \leq f(Y)/e(Y)$.  
```{r}
laccept=U<=f_x(Y)/e_x(Y)
X <- Y[laccept]
#Count and report Acceptance Ratio
length(X)/N
```
The acceptance ratio is about 0.87.   
Step 5: Estimate $E[X^2]$ by taking the mean of recorded X.   
```{r}
mean(X^2)
```
$$\int_{-\infty}^{\infty} \frac{1}{\sqrt{2\pi}} exp(-\frac{x^2}{2})\sqrt{2\pi} exp(-\frac{x^2}{2})x^2dx$$
Where $h(x)=\sqrt{2\pi}exp(-\frac{x^2}{2})x^2$ and $f(x)=\frac{1}{\sqrt{2\pi}}exp(-\frac{x^2}{2})$, $x\sim N(0,1)$   
Thus, $\int_{-\infty}^{\infty} x^2 exp(-x^2)dx=E(h(x))$
```{r}
set.seed(1693)
h_x_2<- function(x){
  (sqrt(2*pi)*exp(-x^2/2)*x^2)
}
n<-10^6
x<-rnorm(n)
mean(sapply(x,h_x_2))
```
### Example 3
The joint uniform density is constant over a given range. In this case, the $f_{x,y}(x,y)=1/4$ in range $[-1,1]\times[-1,1]$ and $f_{x,y}(x,y)=0$ outside the range.  
$$\int_{-1}^{1}\int_{-1}^{1}|x-y|dxdy$$
$$\int_{-1}^{1}\int_{-1}^{1}4|x-y|\frac {1}{4}dxdy$$
Where $h(x,y)=|x-y|$ and $f_{x,y}(x,y)=\frac{1}{4}$
Thus, $\int_{-1}^{1}\int_{-1}^{1}|x-y|dxdy=E(h(x,y))$
```{r}
set.seed(1693)
h_x_3<- function(x,y){
  4*(abs(x-y))
}
n<-10^6
x<-runif(n,-1,1)
y<-runif(n,-1,1)
mean(h_x_3(x,y))
```

